---
title: "Mapping blue helmets"
site: workflowr::wflow_site
output:
  workflowr::wflow_html:
    toc: false
editor_options:
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(dpi=150)
```

##  Visualizing deployment locations

An advantage to the Geo-PKO dataset is that it records the numbers of troops by their specific deployment locations. Therefore, users can quickly visualize *where* active troops are in a mission, be it for a specific mission or a region. Below are some examples of visualization.

To start with, we can take a snapshot of the deployment of all the missions that were active in 2018 in Africa. We start by  pulling the shapefiles for country outlines from the package `rnaturalearth`, and filter for countries in Africa. 

```{r, warning=FALSE, message=FALSE}
library(rnaturalearth)
library(rnaturalearthdata)
library(sf)

world <- ne_countries(scale = "medium", returnclass = "sf")
library(dplyr)
AFR_sf <- world %>% filter(region_un == "Africa")
```

Did that work? We can plot the shapefiles to find out.

```{r}
library(ggplot2)

ggplot(data=AFR_sf) + geom_sf()
```

Next, we need to subset the main dataset to include only entries (1) for the year of 2018 and (2) in the continent Africa, all with (3) our variables of interests. Geo-PKO reports deployment sizes according to the available maps published by the UN. Therefore, to obtain the numbers of troop deployment at the yearly level, we calculate the average number of troops per location over the months recorded. We should end up with something similar to the table below. 


```{r, warning=FALSE, message=FALSE}
library(readr)
library(knitr)
library(kableExtra) 

GeoPKO <- readr::read_csv("data/Geo_PKO_v2.0.csv", col_types = cols(.default="c")) 

GeoPKO2018 <- GeoPKO %>% filter(year==2018) %>%  
  select(mission, year, location, latitude, longitude, no.troops, hq, country) %>%
  mutate_at(vars(latitude, longitude, no.troops), as.numeric) %>% 
  group_by(location, mission, latitude, longitude) %>% 
  mutate(YearlyAverage = round(mean(no.troops, na.rm=TRUE))) %>%
  arrange(desc(hq)) %>% slice(1)

kable(GeoPKO2018, caption = "An extract of the 2018 dataframe") %>% kable_styling() %>% 
  scroll_box(width = "100%", height = "200px")
```

Hold up--the data frame still contains missions that did not take place in Africa. We need to filter this out somehow.

```{r, warning=FALSE, message=FALSE}
library(countrycode)

AFR_list <- codelist %>% filter(continent %in% "Africa")  %>% select(country.name.en) %>% pull()

GeoPKO2018_AFR <- GeoPKO2018 %>% 
  mutate(country=case_when(country=="DRC" ~ "Congo-Kinshasa",
                           TRUE~ as.character(country))) %>% 
  filter(country %in% AFR_list)
```

With that set, we can start plotting the deployment locations and their respective sizes for UN peacekeeping missions in Africa in 2018. 

```{r, warning=FALSE, message=FALSE}
library(ggrepel) #to nudge labels nicely away from geom_point
library(viridis) #for pretty colors
library(ggthemes)

ggplot(data=AFR_sf) + geom_sf() + 
  geom_point(data = GeoPKO2018_AFR, aes(x=longitude, y=latitude, 
                                        size=YearlyAverage, color= YearlyAverage), alpha=.7)+
  scale_size_continuous(name="Average Troop Deployment", range=c(1,12), 
                        breaks=c(0, 100, 300, 500, 1000, 2000, 3000, 4000,5000)) +
  scale_color_viridis(option="cividis", 
                      breaks=c(0, 100, 300, 500, 1000, 2000, 3000, 4000,5000), 
                      name="Average Troop Deployment" ) +
  guides( colour = guide_legend()) +
  geom_point(data = GeoPKO2018_AFR %>% filter(hq==3), aes (x=longitude, y=latitude, shape="HQ"),
             fill = "red", size=2, color="red", alpha=.8)+
  scale_shape_manual(values=c(23), labels=c("HQ"="Mission HQ"), name="")+
  geom_label_repel(data = GeoPKO2018_AFR %>% filter(hq==3), aes(x=longitude, y=latitude, label=mission),
                   min.segment.length = 0, 
                   direction="both",
                   label.size = 0.5,
                   box.padding = 2,
                   size = 3, 
                   fill = alpha(c("white"),0.5),
                   shape=16, 
                   size=2) +
  labs(title ="UN Peacekeeping in Africa - 2018", color='Average Troop Deployment') +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA), 
    panel.background = element_rect(fill = "#f5f5f2", color = NA), 
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(size= 14, hjust=0.01, color = "#4e4d47", margin = margin(b = -0.1, t = 0.8, l = 4, unit = "cm")),
    panel.grid=element_blank(),
    axis.title=element_blank(),
    axis.ticks=element_blank(),
    axis.text=element_blank(),
    legend.key=element_blank()
  )
```

Here is a similar visualization, but this time the color aesthetic for geom_point is mapped to shown country instead.

```{r, warning=FALSE, message=FALSE}
p3 <- ggplot(data=AFR_sf) + geom_sf() +
  geom_point(data=GeoPKO2018_AFR, 
             aes(x=longitude, y=latitude, size=YearlyAverage, color=country), alpha=.4, shape=20)+
  geom_point(data=GeoPKO2018_AFR %>% filter(hq==3), 
             aes(x=longitude, y=latitude), 
             color="black", shape=16, size=2) +
  geom_label_repel(data=GeoPKO2018_AFR %>% filter(hq==3),
                   min.segment.length = 0.2,
                   label.size = 0.5,
                   box.padding = 2,
                   size = 3,
                   fill = alpha(c("white"),0.7),
                   aes(x=longitude, y=latitude, label=mission)) +
  labs(title="UN Peacekeeping Deployment and Mission HQs in Africa, 2018")+
  scale_size(range = c(2, 16))+
  labs(size="Average number of troops\n(continuous scale)",color="Country",shape="HQ")+
  theme(
    plot.background = element_rect(fill = "#f5f5f2", color = NA),
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(size= 14, hjust=0.01, color = "#4e4d47", margin = margin(b = -0.1, t = 0.8, l = 4, unit = "cm")),
    panel.grid=element_blank(),
    axis.title=element_blank(),
    axis.ticks=element_blank(),
    axis.text=element_blank(),
    panel.background=element_blank(),
    legend.key = element_rect(fill = "#f5f5f2", color = NA),
    legend.key.size = unit(1, 'lines')
  )+
  guides(colour=guide_legend(ncol=2,override.aes = list(size=5)),
         size=guide_legend(ncol=2))
p3
```

How has this changed over the period covered by the dataset? An animated graph is great for this purpose. The first step is to prepare a dataframe, much similar to what has been done above for missions taking place in Africa in 2018. First we would calculate the average number of troops that is deployed to a location per mission per year, for every year between 1994 and 2018. 

```{r, warning=FALSE}
AFR_GIF <- GeoPKO %>%   
  mutate(country=case_when(country=="DRC" ~ "Congo-Kinshasa",
                           TRUE~ as.character(country))) %>%
  filter(country %in% AFR_list) %>% 
  select(mission, year, location, latitude, longitude, no.troops, hq) %>%
  mutate_at(vars(latitude, longitude), as.numeric) %>% 
  group_by_at(vars(-no.troops)) %>%
  summarise(ave.no.troops = round(mean(as.numeric(no.troops), na.rm=TRUE)))
```

The animated graph is built on the above code for static graphics, using the cool package `gganimate`.

```{r, animatedgraph, warning=FALSE, dev="png", interval=0.8}
library(gganimate)

# Transforming the "year" variable into a discrete variable.
AFR_GIF$year <- as.factor(AFR_GIF$year)

ggplot(AFR_sf) + geom_sf() + 
  geom_point(data = AFR_GIF, aes(x=longitude, y=latitude, 
                                 size= ave.no.troops, color= ave.no.troops, group=year), alpha=.7)+
  scale_size_continuous(name="Average Troop Deployment", range=c(1,12), 
                        breaks=c(0, 100, 300, 500, 1000, 2000, 3000, 4000,5000)) +
  scale_color_viridis(option="cividis", 
                      breaks=c(0, 100, 300, 500, 1000, 2000, 3000, 4000,5000),
                      name="Average Troop Deployment" ) +
  guides(colour = guide_legend()) +
  theme(
    text = element_text(color = "#22211d"),
    plot.background = element_rect(fill = "#f5f5f2", color = NA), 
    panel.background = element_rect(fill = "#f5f5f2", color = NA), 
    legend.background = element_rect(fill = "#f5f5f2", color = NA),
    plot.title = element_text(size= 14, hjust=0.01, color = "#4e4d47", margin = margin(b = -0.1, t = 0.8, l = 4, unit = "cm")),
    panel.grid=element_blank(),
    axis.text=element_blank(),
    axis.ticks=element_blank(),
    axis.title=element_blank(),
    legend.key=element_blank(),
    plot.caption=element_text(hjust=1, face="italic"))+
  transition_states(states=year, transition_length = 3, state_length=3)+
  labs(title="UN Peacekeeping in Armed Conflicts in Africa: {closest_state}",
       caption="Source: The Geo-PKO dataset v2.0")+
  ease_aes()

#run the following command to save the plot
#anim_save("animatedUNPKO.gif", p4)

```

With the release of version 2.0, the Geo-PKO dataset has extended its coverage to missions taking place globally between 1994 and 2019. The above examples are but a few ways through which users can explore and exploit this rich dataset. 

The following code shows how to make an animate plot with the global data by year, using the package "animate".
```{r, animatedgraph_2, warning=FALSE, dev="png", interval=0.8}
library(tidyr)

#Create a new dataframe keeping all missions and regions
Global_df <- GeoPKO %>% 
  select(mission, year, location, latitude, longitude, no.troops) %>%
  mutate_at(vars(latitude, longitude, no.troops), as.numeric) %>% 
  group_by(mission, year, location) %>%
  mutate(ave.no.troops = as.integer(mean(no.troops, na.rm=TRUE))) %>% 
  select(-no.troops) %>% distinct() %>% drop_na(ave.no.troops) %>% filter(ave.no.troops>0)

#Make the year variable discrete
Global_df$year <- as.factor(Global_df$year)

#Make the plot
ggplot() +     
  borders("world",  xlim = c(-130, 140), ylim = c(-50, 50), colour = "gray85", fill = "gray80") + 
  theme_void()+
  geom_point(data = Global_df, aes(x=longitude, y=latitude, size= ave.no.troops, color= ave.no.troops, group=year), alpha=.7)+
  scale_size_continuous(name="Average Troop\nDeployment", range=c(1,12), breaks=c(10, 100,300, 500, 1000, 3000, 4500, 6000, 7000)) +
  scale_color_viridis(option="viridis", breaks=c(10, 100, 300, 500, 1000, 3000, 4500, 6000, 7000), name="Average Troop\nDeployment" ) +
  guides(colour = guide_legend()) + 
  theme(text = element_text(color = "#22211d"),
        legend.position =  c(0.9,0.4),
        plot.background = element_rect(fill = "#f5f5f2", color = NA), 
        panel.background = element_rect(fill = "#f5f5f2", color = NA), 
        legend.background = element_rect(fill = NA, color = NA),
        legend.key = element_rect(fill = NA, color = NA),
        plot.title = element_text(size= 14, hjust=0.01, color = "#4e4d47", margin = margin(b = -0.1, t = 0.8, l = 4, unit = "cm")),
        #        plot.margin=margin(1, 1, 1, 1, "cm"),
        panel.grid=element_blank(),
        axis.text=element_blank(),
        axis.ticks=element_blank(),
        axis.title=element_blank(),
        plot.caption=element_text(hjust=1, face="italic")) +
 transition_states(states = year, transition_length = 3, state_length=3)+
 labs(title="UN Peacekeeping around the world: {closest_state}",
      color="Average Deployment Size",
      caption="Source: The Geo-PKO dataset v2.0")+
 ease_aes()

#Animate the plot
#animate(global_p, height = 600, width =800, fps = 4, res=120)
##To save the plot run the following line of code
#anim_save("Animated_GeoPKO2.0.gif")
```
